Using simulation to optimize the sampling design of fisheries-independent trawl surveys

### 2018-07-06 analysis/paper/sim_survey_paper.Rmd
Discovered and fixed issue with age-year-space covariance

The objective of this paper is to use a realistic simulation to evaluate the efficacy of various sampling protocol. First we simulated a spatially autocorrelated target population, and then we simulated surveys over this field. Set density was varied across surveys, as was length and age sampling efforts. Estimates of abundance at age were then obtained using a standard stratified approach [@Smith1981] and deviation from the truth was assessed. Given previous findings, we predict that the greatest gains in precision will come from increasing the number of hauls; increased sampling of correlated variables (length and age) on a set by set basis will be relatively ineffective. The bottom trawl survey of cod (*Gadus morhua*) in NAFO division 3Ps, Newfoundland, was used as a case study for this analysis.

@@

- Feedback and advice: Alejandro Buren, Mariano Koen-Alonso, Karen Dwyer, Joanne Morgan, Geoff Evans, Don Power, Brian Healey, Martha Robertson, Danny Ings, Don Stansbury, Mark Simpson, Derek Osborne, Peter Upward, Brad Squires, Dwayne Pittman, Paul Higdon
- Dave Cote for a review of a pervious version of this manuscript

>>>>

The objective of this paper is to use a realistic simulation to evaluate the efficacy of various sampling protocol. First we simulated a spatially autocorrelated target population, and then we simulated surveys over this field. Set density was varied across surveys, as was length and age sampling efforts. Estimates of abundance at age were then obtained using a standard stratified approach [@Smith1981] and deviation from the truth was assessed. Given previous findings, we predict that the greatest gains in precision will come from increasing the number of hauls; increased sampling of correlated variables (length and age) on a set by set basis will be relatively ineffective. The bottom trawl survey of cod (*Gadus morhua*) in NAFO division 3Ps, Newfoundland, was used as a case study for this analysis. While our case study is specific, we also hope that the approach presented here will provide a framework for optimizing sampling protocol for an array of species with different life-histories and distributions. To facilitate future work and collaborations, we have developed an R package for conducting the tests presented in this paper (https://github.com/PaulRegular/SimSurvey).

# Methods

The simulation presented here starts with a common cohort model where the abundance at age $a$ in year $y$ ($N_{a,y}$) equals the abundance of that cohort in the previous year multiplied by the survival rate, which is expressed in terms of total mortality ($Z_{a,y}$):

$$N_{a,y} = N_{a-1,y-1} e^{-Z_{a-1,y-1}}$$
Here, numbers-at-age in the first year are filled via exponential decay, $N_{a,1} = N_{a-1,1} e^{-Z_{a-1,1}}$, numbers at age 1 (i.e. recruits) are filled using a random walk, $log(N_{1,y}) = log(N_{1,y-1}) + \epsilon_{y-1}$, and total mortality is set to a baseline level plus process error, $log(Z_{a,y}) = log(z) + \delta_{a,y}$. The error around the recruitment process was simulated using the normal distribution, $\epsilon_{y} \sim N(0, \sigma_{r}^{2})$, and the process error was simulated using the covariance structure described in Cadigan [-@Cadigan2016], $\delta_{a,y} \sim N(0, \Sigma_{a,y})$. The covariance across ages and years is controled by a process error variance parameter ($\sigma_{\delta}^2$) along with age and year correlation parameters ($\varphi_{\delta, age}$ and $\varphi_{\delta, year}$, respectively). This structure allows for autocorrelation in process errors across ages and years (i.e. total mortality can be made to be more similar for fish that are closer together in age and time). Overall, this  formulation allows for the simulation of an array of population dynamics that one might expect to observe.

The next step in the simulation is to distribute the abundance-at-age matrix simulated using the cohort model throughout a spatial field. Here, a grid of $s$ locations was generated with a resolution of $r$, a depth gradient, $S$ depth-based strata and $D$ divisions. The depth gradient was set-up to have a sigmoid shape with a depth range of $[d_{min}, d_{max}]$, shelf depth of $d_{shelf}$ and a shelf width of $w_{shelf}$. The simulated population is distributed through the grid by simulating spatial-temporal noise controled by a parabolic relationship with depth and covariance between ages, years and space

$$\begin{aligned}log(N_{a,y,s}) &= log(N_{a,y}) + \eta_{a,y,s} \\ \eta_{a,y,s} &= \frac{(d_s - \mu_{d}) ^ 2}{2 \sigma_d^2} + \xi_{a,y,s} \end{aligned}$$

Where $d_s$ is the depth in a specific cell of the grid, $\mu_d$ is the mean depth where abundance is typically highest and $\sigma_d$ controls the width or disperson of abundance around the mean. This helps simulate depth preferences of praticular fish. Residual noise $\xi_{a,y,s}$, which is assumed to follow $\xi_{a,y,s} \sim N(0, \Sigma_{a,y,s})$, is added to this depth relationship using a combination of Matérn covariance, to control the level of spatial aggregation, and the age-year covariance described in Cadigan [-@Cadigan2016], to control the level of similarity in distributions across ages and years. Matérn covariance was controled by a variance ($\sigma_s^2$), smoothing ($\lambda$) and scaling parameter ($\kappa$) [@Blangiardo2015]. Like the process error, age and year covariance was controled by a variance parameter ($\sigma_{\xi}^2$) and two correlation parameters ($\varphi_{\xi, age}$ and $\varphi_{\xi, year}$). In short, this formulation allows control of depth preferences, the level of spatial aggregation and the amount of age and year specific clustering. Note that $\eta_{a,y,s}$ is forced to sum to one in each year to ensure that the total population of each age for each year through the grid equals the number simulated by the cohort model. A more detailed description of the space-age-year covariance is included in Appendix 1.

The final step in the simulation is to conduct a stratified random survey over the age-year-space array generated using formula described above.

@@

- Feedback and advice: Alejandro Buren, Mariano Koen-Alonso, Karen Dwyer, Joanne Morgan, Geoff Evans, Don Power, Brian Healey, Martha Robertson, Danny Ings, Don Stansbury, Mark Simpson, Derek Osborne, Peter Upward, Brad Squires, Dwayne Pittman, Paul Higdon
- Dave Cote for a review of a previous version of this manuscript

# Appendix 1: age-year-space covariance

The simulation applied in this paper was set-up to control covariance across ages, years and space. Matern covariance was used to control the level of spatial aggregation:

$$\Sigma_{s} = Cov(s_i, s_j) = \sigma_s^2 \frac{2^{1-\lambda}}{\Gamma(\lambda)} (\kappa \left |s_i - s_j \right|) K_{\lambda}(\kappa \left |s_i - s_j \right|)$$

where $\left |s_i - s_j \right|$ is the Euclidean distance between two locations, $\sigma_s^2$ is the marginal variance, $\Gamma$ is the gamma function, $K_{\lambda}$ denotes the modified Bessel function of the second kind, and $\lambda$ and $\kappa$ are parameters of the covariance [@Blangiardo2015]. This spatial covariance was used in conjunction with covariance described in Cadigan [-@Cadigan2016] to simulate error across ages, years and space:

$$\xi_{a,y,s} = \left\{\begin{matrix}
 0 + v_{a,y} + s_{s} & i = 1, j = 1 & \\ 
 &  & 
\end{matrix}\right.$$
